Grounding the interaction: anchoring situated discourse in everyday human-robot interaction

### 2011-07-16 paper_knowledge.tex
Raquel's changes

\subsection{Situated speech acts}
\label{intro_example}

@@

While this kind of interaction should hopefully sound quite familiar in a
foreseeable future, our robots are not yet quite up to the task, neither
regarding natural language understanding nor plan-making and manipulation.

To be combined together, those abilities require an unambiguous and shared
representation of concepts (objects, agents, actions...) underlying the
interaction: what are the prerequisites for such a sentence --``Jido,
give me that''-- to be understood by the robot, correctly interpreted in the
spatial context of the interaction, and eventually transformed into an action?

Austin~\cite{Austin1962} would have at first glance analyzed this kind of
sentence as a \emph{speech act}, comprising of \emph{locutionary},
\emph{illocutionary} and possibly \emph{perlocutionary} acts. First, we want to
understand the direct meaning of the sentence (\emph{locutionary act}): we must
acquire the sentence, convert it into a useful syntactic form (probably through
speech recognition), and understand the semantics of the sentence, \ie What is
referred by ``\textit{Jido}''? What is ``\textit{give}''? What is
``\textit{me}''? And ``\textit{that}''?

Working in a situated context, we want furthermore to \emph{resolve} these
semantics atoms, \ie ground them in the sensory-motor space of the robot. For
instance, ``\textit{this}'' is a demonstrative pronoun that refers in this
context to the object the human is focusing on (whatever \textit{focusing}
means). In this example Thomas is looking at something, which is a possible
cue. But it could as well be pointing at something or making reference to some
previously mentioned concept.

@@

Second, the \emph{illocutionary force}, \ie the \emph{intent} of the utterance
as thought by the agent, must be extracted and understood. In our example,
Thomas obviously wants an action to be performed by the robot. The action
parametrization is conveyed by the semantics attached to the words and the
grammatical structure of the sentence. In our example, the type of action is
given by the verb ``\textit{give}''. Assuming the robot has some procedural
knowledge attached to this symbol, the action type can be considered as
grounded for the robot. We can as well understand that the recipient of the
action is the human, the performer is the robot itself, and the object acted
upon is the tape. These are the basic \emph{thematic roles}~\cite{Gruber1965}
that can be extracted from the sentence that allow for fully grounding the action
\footnote{The third class of speech acts, the \emph{perlocutionary acts}, deal
with implicit speaker's expectations. We will not consider it further in this
article}.

%Finally, the implicit speaker's expectation can sometimes be conveyed --the
%\emph{perlocutionary act}-- and should be taken into account. If Thomas would
%have said instead ``Jido, this box is too heavy for me'', he may implicitly
%expect the robot to memorize the fact that the box's weight is over his
%capabilities, and it may as well imply that Thomas is requiring help to carry
%the box.

@@

Extracting these speech acts and turning them into content processable by the
robot is a difficult challenge in the general case. We base our approach on
three distinct, inter-related cognitive functions.
Figure~\ref{fig|overall-process} illustrates the general context, while
Figure~\ref{fig|main-components} summarizes the interactions between the three
main components:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Technical analysis}
\label{examples}

In order to better understand the overall process we next describe the
different steps of the approach based on three examples. In these examples, we
assume that some initial facts are present in the knowledge base, both in the
robot's own model and in the human's model.  Since the robot tries to ground a
human utterance, all queries are sent to the human model in order to interpret
it from the human perspective.

@@

To resolve the nominal group \emph{The yellow banana} a set of partial
statements that describe the concept is generated based on the grammatical
parsing of the sentence (algorithm~\ref{algo|Resolution},
\emph{(\ref{res.desc})}). In the example, a banana (\stmt{?obj type Banana})
that is yellow (\stmt{?obj hasColor yellow})\footnote{Predicates like
\concept{hasColor} or \concept{hasSize} that bind \concept{banana\_01} to
adjectives are extracted from a predefined database of $[Predicate \rightarrow
AdjectiveCategory]$, and falls back on the generic \concept{hasFeature}
predicate if the adjective is not known.}.  Based on these partial statements a
query is sent to the ontology server to retrieve possible instances that match
the description (algorithm~\ref{algo|Resolution}, \emph{(\ref{res.onto})}).

@@

This last example (Figure~\ref{dialog|ex3}) shows the resolution of ambiguous
concepts. In this case the user refers to ``the banana'' while two instances of 
the \concept{Banana} class exist in the ontology. The robot needs to find out
to which instance the user is actually referring to. To this end,
disambiguation routines (algorithm~\ref{algo|Resolution},
\emph{(\ref{st.discrimination})}, see \cite{Ros2010b} for details of the routines)
find differences between the instances (in the example, one banana is yellow
while the other one is green) and build a sentence through the
\emph{verbalization} module to ask the user a closed question that will
help clarify the ambiguity: ``Is it yellow or green?'' The user's answer is
parsed and merged with the previous sentence. The resulting, augmented,
sentence (``The green banana is good") goes again through all the
interpretation steps. This process is repeated until no ambiguities arise. 
In the example, the \concept{banana\_02} is finally returned.

@@

\subsection{``Moving to London'' scenario}

@@

First Tom enters the room with some of the things they need to pack: a
toolbox and two videos. He leaves one of the videos (``Jido-E'') inside one of
the boxes (the cardboard box), and the other one (``Lord of the Robots'') on the
table. We next describe how the situation assessment takes place, \ie how the
ontology is updated with the information obtained from the geometric reasoner
SPARK.

The initial information in ORO corresponds to:

\begin{center}
    \begin{tabular}{l}
      \stmt{table type Table}\\
      \stmt{cardBoardBox type Box}\\
      \stmt{toolbox type Box}\\
      \stmt{videoTape1 type VideoTape}\\
      \stmt{videoTape1 label "Lord of the Robots"}\\
      \stmt{videoTape2 type VideoTape}\\
      \stmt{videoTape2 label "Jido-E"}\\
      \stmt{Tom type Human}\\
      \stmt{Jerry type Human}
    \end{tabular}
\end{center}
\vspace{0.10cm}

SPARK detects that there is a cardboard box on the table. It thus sends the fact to ORO:

>>>>

%\subsection{Situated speech acts}
%\label{intro_example}

@@

While this kind of interaction should hopefully sound quite familiar in a
foreseeable future, our robots are not yet quite up to the task, neither
regarding natural language understanding nor plan-making and manipulation.
To be combined together, those abilities require an unambiguous and shared
representation of concepts (objects, agents, actions...) underlying the
interaction: what are the prerequisites for such a sentence --``Jido,
give me that''-- to be understood by the robot, correctly interpreted in the
spatial context of the interaction, and eventually transformed into an action?

The first step is to understand the meaning of the sentence. To this end, we must acquire the sentence, convert it into a useful syntactic form (probably through
speech recognition), and understand the semantics of the sentence, \ie What is
referred by ``\textit{Jido}''? What is ``\textit{give}''? What is
``\textit{me}''? And ``\textit{that}''?

Working in a situated context, we need to \emph{resolve} these
semantics atoms, \ie ground them in the sensory-motor space of the robot. For
instance, ``\textit{this}'' is a demonstrative pronoun that refers in this
context to the object the human is focusing on (whatever \textit{focusing}
means). In this example Thomas is looking at something, which is a possible
cue. But it could as well be pointing at something or making reference to some
previously mentioned concept.

The next step is to extract and understand the \emph{intent} of the utterance
as thought by the agent. In our example, Thomas obviously wants an action to be performed by the robot. The action parametrization is conveyed by the semantics attached to the words and the grammatical structure of the sentence. In our example, the type of action is
given by the verb ``\textit{give}''. Assuming the robot has some procedural
knowledge attached to this symbol, the action type can be considered as
grounded for the robot. We can as well understand that the recipient of the
action is the human, the performer is the robot itself, and the object acted
upon is the tape. These are the basic \emph{thematic roles}~\cite{Gruber1965}
that can be extracted from the sentence that allow for fully grounding the action.\footnote{This analysis has been inspired on the work of Austin et al.~\cite{Austin1962}, where this type of sentences correspond to \emph{speech acts}, comprising of \emph{locutionary act} (the meaning), \emph{illocutionary} (the intent) and possibly \emph{perlocutionary acts} (implicit speaker's expectation).}

%\footnote{The third class of speech acts, the \emph{perlocutionary acts}, deal
%with implicit speaker's expectations. We will not consider it further in this
%article}.
%Austin~\cite{Austin1962} would have at first glance analyzed this kind of
%sentence as a \emph{speech act}, comprising of \emph{locutionary},
%\emph{illocutionary} and possibly \emph{perlocutionary} acts. 
%Firstly, we want to understand the direct meaning of the sentence (\emph{locutionary act}): 
%Second, the \emph{illocutionary force}, \ie the \emph{intent} of the utterance
%as thought by the agent, must be extracted and understood. 
%Finally, the implicit speaker's expectation can sometimes be conveyed --the
%\emph{perlocutionary act}-- and should be taken into account. If Thomas would
%have said instead ``Jido, this box is too heavy for me'', he may implicitly
%expect the robot to memorize the fact that the box's weight is over his
%capabilities, and it may as well imply that Thomas is requiring help to carry
%the box.

@@

@@

In this work we propose an approach aiming at grounding the interaction with users, verbal and non-verbal, following the steps described above. To this end, we have developed three distinct, inter-related cognitive components, which we believe fulfill the minimum requisites for situated human interaction. Figure~\ref{fig|overall-process} illustrates the general context, while
Figure~\ref{fig|main-components} summarizes the interactions between these three
main components.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Technical analysis}
%\label{examples}

In order to better understand the overall process the {\sc Dialogs} module and its relation with ORO, we next describe the different steps of the approach based on three examples. In these examples, we
assume that some initial facts are present in the knowledge base, both in the
robot's own model and in the human's model.  Since the robot tries to ground a
human utterance, all queries are sent to the human model in order to interpret
it from the human perspective.

@@

To resolve the nominal group \emph{The yellow banana} a set of partial
statements that describe the concept is generated based on the grammatical
parsing of the sentence (algorithm~\ref{algo|Resolution}(\ref{res.desc})). In the example, a banana (\stmt{?obj type Banana})
that is yellow (\stmt{?obj hasColor yellow})\footnote{Predicates like
\concept{hasColor} or \concept{hasSize} that bind \concept{banana\_01} to
adjectives are extracted from a predefined database of $[Predicate \rightarrow
AdjectiveCategory]$, and falls back on the generic \concept{hasFeature}
predicate if the adjective is not known.}.  Based on these partial statements a
query is sent to the ontology server to retrieve possible instances that match
the description (algorithm~\ref{algo|Resolution}(\ref{res.onto})).

@@

This last example (Figure~\ref{dialog|ex3}) shows the resolution of ambiguous
concepts. In this case the user refers to ``the banana'' while two instances of 
the \concept{Banana} class exist in the ontology. The robot needs to find out
to which instance the user is actually referring to. To this end,
disambiguation routines (algorithm~\ref{algo|Resolution}(\ref{st.discrimination}), see \cite{Ros2010b} for details of the routines)
find differences between the instances (in the example, one banana is yellow
while the other one is green) and build a sentence through the
\emph{verbalization} module to ask the user a closed question that will
help clarify the ambiguity: ``Is it yellow or green?'' The user's answer is
parsed and merged with the previous sentence. The resulting, augmented,
sentence (``The green banana is good") goes again through all the
interpretation steps. This process is repeated until no ambiguities arise. 
In the example, the \concept{banana\_02} is finally returned.

@@

\subsection{\emph{Moving to London} scenario}

@@

First Tom enters the room with some of the things they need to pack: a
toolbox and two videos. He leaves one of the videos (``Jido-E'') inside one of
the boxes (the cardboard box), and the other one (``Lord of the Robots'') on the
table. We next describe how the situation assessment takes place, \ie how the
ontology is updated with the information obtained from the geometric reasoner
SPARK.
The initial information in ORO corresponds to:

\begin{center}
    \begin{tabular}{l}
      \stmt{table type Table}\\
      \stmt{cardBoardBox type Box}\\
      \stmt{toolbox type Box}\\
      \stmt{videoTape1 type VideoTape}\\
      \stmt{videoTape1 label "Lord of the Robots"}\\
      \stmt{videoTape2 type VideoTape}\\
      \stmt{videoTape2 label "Jido-E"}\\
      \stmt{Tom type Human}\\
      \stmt{Jerry type Human}
    \end{tabular}
\end{center}
\vspace{0.10cm}
	SPARK detects that there is a cardboard box on the table. It thus sends the fact to ORO:
